---
title: "Penguin Anova"
author: "Fernanda Montero"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read-in penguins.csv data
```{r, include = FALSE}
library(tidyverse)
library(dplyr)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring

New funtions: with = table for easy counting over multipe factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>% 
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           ) 
           ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")


```

- Exploratory graphs
- Summary statistics
- Levene's tests for equal variances 
- one-way ANOVA
- post-hoc testing

```{r}

hist <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +   # data does look normally distributed and 
  facet_wrap(~ common_name)

hist

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +  
  facet_wrap(~ common_name)

qqs

# Levene's test for equal variances (> 2 groups... if you have two groups, use the F-test for equal variances)

#We'll use the funtion leveneTest form the 'car' package

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

#We retain the null hypothesis of equal variances (p=0.06)

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances

```

3. One-way ANOVA

Single factor: common_name (species)
Number of levels in the factor: 3 (Chistrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significant difference in mean flipper length (mm) for female Chisptrap, Gentoo and Adelie penguins?

```{r}

#H0: Mean flipper lengths across all species are equal
#HA: At LEAST two means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

#At least two samples were taken from populaions with different means. Which ones are different? All are different from each other? Or something else?

#Post-hoc testing using Turkey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

```

Mean flipper lenghts for Adelie, Gentoo, and Chinstrap penguins are all significantly different (F(df)=statistic, p < 0.001, $\alpha\$ = 0.05 with post-hoc Tukey's HSD, $\alpha = 0.05$.



